---
title: "Exercise 3"
subtitle: "R Markdown SQL vs dplyr backend"
output: html_document
---

You can query your database directly from R Markdown using traditional SQL syntax. Make sure you use RStudio version 1.0 or higher and use the SQL engine in the code chunk as described [here](http://rmarkdown.rstudio.com/authoring_knitr_engines.html#sql).

You can also query your database directly from R Markdown using dplyr syntax. Dplyr has backend connectivity for a variety of databases that will translate your R code into SQL. See [here](https://cran.r-project.org/web/packages/dplyr/vignettes/databases.html) for more information.

In this exercise you will use both SQL and dplyr to connect to a SQLite database, collect the data into R, and then analyze the data with familiar R tools. 

### Intro

The load the required packages. Set up a SQL light database and populate it with data from the `nycflights13` R package. This step only needs to be run once. 

```{r}
library(dplyr)
library(ggplot2)
library(highcharter)
library(DBI)
library(nycflights13)
```

```{r setup, eval = FALSE, include = FALSE}
file.remove("my_db.sqlite3")
my_db <- dplyr::src_sqlite("my_db.sqlite3", create = TRUE)
dplyr::copy_to(my_db, nycflights13::flights, "flights", temporary = FALSE)
dplyr::copy_to(my_db, nycflights13::airlines, "airline", temporary = FALSE)
```

***

### Compare SQL to dplyr syntax 

Use SQL and dplyr syntax to query the database. Create a query with the following modifications on the flights table:

* Select only these columns: `carrier`, `dep_delay`, `distance`
* Create a new column `air_time_hours` that is the hourly representation of `air_time`
* Subset the data from May 17
* Subset the data for only these carriers: `UA`, `WN`, `AA`, `DL`

Create the queries for both SQL and dplyr. Notice that the output of the queries are identical. Render the SQL translation from dplyr. Notice that the translated syntax looks slightly different but accomplishes the same result.

#### SQL

```{r}
db <- DBI::dbConnect(RSQLite::SQLite(), dbname = "my_db.sqlite3")
```

```{sql, connection=db}
SELECT carrier, dep_delay, distance, air_time, air_time / 60 as air_time_hours
FROM flights 
WHERE month = 5 AND day = 17 AND carrier IN ('UA', 'WN', 'AA', 'DL')
```

#### dplyr

```{r}
my_db <- src_sqlite("my_db.sqlite3")
airline <- tbl(my_db, "airline")
flights <- tbl(my_db, "flights")
```

```{r}
flights %>%
  filter(month == 5, day == 17, carrier %in% c('UA', 'WN', 'AA', 'DL')) %>%
  mutate(air_time_hours = air_time / 60) %>%
  select(carrier, dep_delay, distance, air_time, air_time_hours)
```


```{r}
sql_render(
  flights %>%
  filter(month == 5, day == 17, carrier %in% c('UA', 'WN', 'AA', 'DL')) %>%
  mutate(air_time_hours = air_time / 60) %>%
  select(carrier, dep_delay, air_time, distance)
)
```

***

### Collecting

After you run your queries you may want to bring your data back into R for further analysis. To bring your data into R, use the `output.var` or the `collect` statements in SQL and dplyr respectively. Run the queries below. Compare the output. Notice that they returned the same contents. Rename the output to `carrierhours`.

The queries below have the following these additional modifications:

* Join the carrier table
* Rename `name` to `carrier_name`
* Sort the data by carrier

```{sql, connection=db, output.var="carrierhours_sql", include=TRUE}
SELECT b.name as carrier_name, a.*
FROM (
    SELECT carrier, dep_delay, air_time, distance, air_time / 60 as air_time_hours
    FROM flights 
    WHERE month = 5 AND day = 17 AND carrier IN ('UA', 'WN', 'AA', 'DL')
    ) a
LEFT JOIN airline b USING (carrier)
ORDER BY carrier
```

```{r, include=TRUE}
carrierhours_dplyr <- flights %>%
  filter(month == 5, day == 17, carrier %in% c('UA', 'WN', 'AA', 'DL')) %>%
  left_join(airline, by = "carrier") %>%
  select(name, carrier, dep_delay, air_time, distance) %>%
  arrange(carrier) %>%
  mutate(air_time_hours = air_time / 60) %>%
  rename(carrier_name = name) %>%
  collect
```

```{r}
carrierhours_sql <- as.tbl(carrierhours_sql)
identical(carrierhours_dplyr, carrierhours_sql)
carrierhours <- carrierhours_dplyr
```

***

### Analysis

You can use the same dplyr syntax to manipulate your data in R that you used to query your database. The code below uses dplyr syntax to calculate the average departure dealy, distance, and air time by carrier. The code also runs a predictive model on the collected data. Finally, you can visualize your data with the `ggplot2` and `highcharter` packages.

```{r}
# Carrier averages
carrierhours %>%
  group_by(carrier_name) %>%
  summarize(n = n(), 
            dep_delay = mean(dep_delay), 
            distance = mean(distance), 
            air_time_hours = mean(air_time_hours))

# Predictive model
lm1 <- lm(dep_delay ~ air_time_hours + distance + carrier_name, data = carrierhours)
summary(lm1)
```

```{r}
hcboxplot(carrierhours$air_time, carrierhours$carrier_name, color = "darkgrey", name = "carrier")
ggplot(carrierhours, aes(carrier_name, air_time_hours)) + geom_boxplot()
```

***

### Exercise

In this exercise you will query the same database with both SQL and dplyr. Your query will contain the average departure delays and arrival delays for flights between "JFK" and "SFO" by month. Specifically, your query should have the following modifications on the `fligths` table:

* Filter the data to include only flights from "JFK" to "SFO"
* Group by month
* Summarize your so your output calculates the med `dep_delay` and the mean `arr_delay`

Write your data in dplyr and sql. Collect your data. Then create a scatterplot that shows the relationship between mean departure delays and arrival delays.

Hint: You can use `sql_render` to generate the SQL statement from you dplyr code.

Post your completed R Markdown document on RStudio connect and email Nathan.

### Solution

```{r}
ex1_dplyr <- flights %>%
  filter(origin == "JFK" & dest == "SFO") %>%
  group_by(month) %>%
  summarize(dep_delay = mean(dep_delay), arr_delay = mean(arr_delay)) %>%
  collect

qplot(dep_delay, arr_delay, data = ex1_dplyr, geom = "point")
```

```{r}
sql_render(flights %>%
  filter(origin == "JFK" & dest == "SFO") %>%
  group_by(month) %>%
  summarize(dep_delay = mean(dep_delay), arr_delay = mean(arr_delay))
)
```

```{sql, connection=db, output.var="ex1_sql"}
SELECT `month`, AVG(`dep_delay`) AS `dep_delay`, AVG(`arr_delay`) AS `arr_delay`
FROM (SELECT *
FROM `flights`
WHERE (`origin` = 'JFK' AND `dest` = 'SFO'))
GROUP BY `month`
```

```{r}
qplot(dep_delay, arr_delay, data = ex1_sql, geom = "point")
```

