---
output:
  html_notebook: null
  html_document: null
---

```{r setup}
library(ggplot2)
library(plotly)
```


## In-Line Preview

You can now execute R code in-line and see a preview `r Sys.Date()` (Just place the cursor on the in-line code and hit CMD (ctrl) + Enter)

You can preivew LaTex in the same way: $y = \frac{1}{x}$

## Gallery View and Satellite Chunks

If a code chunk generates more than one output they are displayed across thumbnails:

```{r}
ggplot(mtcars, aes(x=hp, y=mpg)) + geom_point() + geom_smooth()
p <- ggplot(mtcars, aes(x=hp, y=mpg)) + geom_point(aes(colour = as.factor(cyl))) + geom_smooth()
p
ggplotly(p)
```

## Tabular Output

By default, printing tabular results (like `data.frame`) will create an interactive table:

```{r}
iris
```


## Multiple Language Support

You can insert chunks for different language types through the Insert Chunk UI

The chunk settings for different languages has been improved and includes defining the engine path. You can run different versions of languages by defining the install location in `engine.path`:


```{python, engine.path="/usr/bin/python2.7"}
import sys
print(sys.version)
```

```{python, engine.path="/usr/bin/python3"}
import sys
print(sys.version)
```

One of the newly supported languages is SQL:

First, create a db in R:

```{r}
d <- datasetsDb()
```

```{sql connection=d}
SELECT * FROM iris
```
